---
title: "Exercise package building"
author: "Sophie C. Schmidt"
date: "7 6 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Create your package!

```{r}
usethis::create_package("/path/to/myAMAZINGpackage")
```

R will tell you, what it does, it should look something like this: 
✓ Creating '/path/to/myAMAZINGpackage/'
✓ Setting active project to '/path/to/myAMAZINGpackage'
✓ Creating 'R/'
✓ Writing 'DESCRIPTION'
... tells you what it writes in there
✓ Writing 'NAMESPACE'
✓ Writing 'myAMAZINGpackage.Rproj'
✓ Adding '^myAMAZINGpackage\\.Rproj$' to '.Rbuildignore'
✓ Adding '.Rproj.user' to '.gitignore'
✓ Adding '^\\.Rproj\\.user$' to '.Rbuildignore'
✓ Opening '/path/to/myAMAZINGpackage' in new RStudio session
✓ Setting active project to '<no active project>'

So you now have a new Rstudio project and the bare bones of a package.

In Rstudio Cloud now make sure your build tools are up for package building:
- upper horizontal menu -> Build -> Configure Build Tools -> in drop down menu "project build tools" choose *package*
- a new tab on the upper right hand corner (next to environment) should appear